From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-07-28 WOLFHPC15/article/article.tex
wolfhpc15 21_07
wolfhpc15 23_07
wolfhpc15 27_07
minor fixes
fix compiler 28_07

\documentclass{acm_proc_article-sp}
\usepackage{tikz}
\usepackage{listings}
%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    smallcp/.style={
    rectangle,rounded corners=3pt,draw=black,font=\tiny,
    },
    resource/.style={
    rectangle,rounded corners=3pt, dashed,draw=black,font=\tiny,
    },
    connection/.style={
    decoration={ markings,
      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
      },
    postaction={decorate}
    },
    bconnection/.style={
    decoration={ markings,
      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
      },
    postaction={decorate}
    },
    mpiconn/.style={
    decoration={ markings,
      mark=at position 0.99 with {\node[fill=black,draw=black,rectangle,,inner sep=2pt](mpi){};}
      },
    postaction={decorate}
    },
}
%-------------

@@

\title{A component-based DSL for multi-stencil programs}

\numberofauthors{2} 
\author{
% 1st. author
\alignauthor
Christian Perez\\
       \affaddr{INRIA}\\
       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{christian.perez@inria.fr}
% 2nd. author
\alignauthor
Helene Coullon\\
       \affaddr{INRIA}\\
       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{helene.coullon@inria.fr}
}

@@

\keywords{Component models, domain specific languages, multi-stencil, numerical simulations} % NOT required for Proceedings

%----------------------------------------
\section{Introduction}
\label{sect:intro}
\input{introduction.tex}
%----------------------------------------
\section{Underlying concepts}
\label{sect:concept}
\input{concept.tex}
%----------------------------------------
\section{The Multi-Stencil Component Assembly Compiler}
\label{sect:mscac}
\input{mscac.tex}
%----------------------------------------
\section{MS Language and MSC compiler}
\label{sect:msmsc}
\input{ms_msc.tex}
%----------------------------------------
\section{Control components and component assembly}
\label{sect:component}
\input{component.tex}
%----------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%----------------------------------------
\section{Related Work}
\label{sect:related}
\input{related.tex}
%----------------------------------------
\section{Conclusion}
\label{sect:conclusion}
\input{conclusion.tex}
%----------------------------------------

>>>>

\documentclass{acm_proc_article-sp}
\usepackage{tikz}
\usepackage{listings}
\usepackage{url}
\usepackage{cite}
%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    smallcp/.style={
    rectangle,rounded corners=3pt,draw=black,font=\tiny,
    },
    resource/.style={
    rectangle,rounded corners=3pt, dashed,draw=black,font=\tiny,
    },
    connection/.style={
    decoration={ markings,
      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
      },
    postaction={decorate}
    },
    bconnection/.style={
    decoration={ markings,
      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
      },
    postaction={decorate}
    },
    mpiconn/.style={
    decoration={ markings,
      mark=at position 0.99 with {\node[fill=black,draw=black,rectangle,,inner sep=2pt](mpi){};}
      },
    postaction={decorate}
    },
}
%-------------
\newenvironment{mydef}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

@@

\title{From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study}

\numberofauthors{2} 
\author{
% 1st. author
\alignauthor
Julien Bigot\\
       \affaddr{CEA, Maison de la Simulation}\\
       \affaddr{USR 3441}\\
       \affaddr{91191 Gif-sur-Yvette Cedex, France}\\
       \email{julien.bigot@cea.fr}
% 1st. author
\alignauthor
Helene Coullon\\
       \affaddr{INRIA}\\
       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{helene.coullon@inria.fr}
% 2nd. author
\and
\alignauthor
Christian Perez\\
       \affaddr{INRIA}\\
       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{christian.perez@inria.fr}
}

@@

%\keywords{Component models, domain specific languages, multi-stencil, numerical simulations} % NOT required for Proceedings

%----------------------------------------
\section{Introduction}
\label{sect:intro}
\input{introduction.tex}
%----------------------------------------
\section{Stencil and multi-stencil}
\label{sect:concept}
\input{concept.tex}
%----------------------------------------
\section{Overview}
\label{sect:mscac}
\input{mscac.tex}
%----------------------------------------
\section{MS Language and MSC compiler}
\label{sect:msmsc}
\input{ms_msc.tex}
%----------------------------------------
\section{Dump to a component assembly}
\label{sect:component}
\input{component.tex}
%----------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%----------------------------------------
\section{Related Work}
\label{sect:related}
\input{related.tex}
%----------------------------------------
\section{Conclusion}
\label{sect:conclusion}
\input{conclusion.tex}
%----------------------------------------

### 2015-07-28 WOLFHPC15/article/component.tex
fix compiler 28_07

@@

\begin{figure*}
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component] (seq0) at (0,0) {$SEQ$};
   \node[provide] (seq0p) [left = 1cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2.5cm of seq0u] {$c_0$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 1cm of sync0] {$c_1^*$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 1cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
   %par1
   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
   \node[use] (par1u) [right = 1cm of par1] {$m$};
  %c2
   \node[component] (c2) [right = 1cm of par1u] {$c_2$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 1cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 1cm of par1] {$c_4^*$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 1cm of c4] {$c_6$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 3cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$c_3$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 1cm of c3] {$c_5$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 1cm of par0] {$c_7$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 1cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 1cm of sync2] {$c_8^*$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

>>>>

%----------------------------------------
\subsection{Component models}
%----------------------------------------
Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities called components. Component models improves code re-use by a separation of concerns, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}. In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.

In the rest of this paper will be used concepts of the component model $L^2C$~\cite{l2c}. $L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world. As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses). 
A \emph{port} in $L^2C$ is an entity embbedded in the component which makes possible an \emph{assembly} of components. An assembly of components is a way to actually connect components together and to produce a complete application, as a set of components and their interactions. A provided service inside a component is associated to a \emph{provide} port, while a required service is associated to a \emph{use} port. It is also possible to group more than one required service into a single port, as a list, called a \emph{use-multiple} port. As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle on the left of the component, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it, both on the right of the component.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$Comp_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=1.5cm of u] (p1) {};
   \node[component,right=1.5cm of p1] (C1) {$Comp_1$};
   \node[use,right=1.5cm of C1] (um) {$m$};

  \path[-]
    (p) edge node {} (C)
    (C.east) edge node {} (u)
    (C1)  edge node {} (um)
    (p1) edge node {} (C1);
\end{tikzpicture}
\caption{Two components, one with a provide, use and use-multiple ports, the second with a single provide port}
\label{fig:ports}
\end{center}
\end{figure}

In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}. As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$Comp_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$Comp_1$};
   \node[use,right=1.5 of C1] (um) {$m$};

  \path[-]
    (p) edge node {} (C)
    (C) edge node {} (u)
    (C1)  edge node {} (um)
    (u) edge node {} (C1);
\end{tikzpicture}
\caption{Component assembly of Figure~\ref{fig:ports}}
\label{fig:assembly}
\end{center}
\end{figure}

With its minimal set of ports, $L^2C$ is a light component model which does not introduce overheads at runtime, and which results in a HPC-oriented component model. Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a distribution of component on distributed architectures is managed. Actually, in CCA distant components (on different processors or machines) have to be managed manually inside components using MPI~\cite{Graham2009MSE} (Message Passing Interfaces), while $L^2C$ proposes a specific MPI port to link distant components and ease their communications. Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside the assembly, which increase separation of concerns and code reuse for parallel applications.

@@

\begin{figure*}
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component] (seq0) at (0,0) {$SEQ$};
   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2cm of seq0u] {$c_0$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 0.8cm of sync0] {$c_1^*$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
   %par1
   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
   \node[use] (par1u) [right = 1cm of par1] {$m$};
  %c2
   \node[component] (c2) [right = 1cm of par1u] {$c_2$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 0.8cm of par1] {$c_4^*$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 0.8cm of c4] {$c_6$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 2.5cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$c_3$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 0.8cm of c3] {$c_5$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 0.8cm of par0] {$c_7$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 0.8cm of sync2] {$c_8^*$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

### 2015-07-28 WOLFHPC15/article/concept.tex
fix compiler 28_07

%----------------------------------------
\subsection{Parallelization techniques}
\label{sect:parallel}
%----------------------------------------
Three classical parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism. Those parallelization techniques are independent from the actual parallel hardware used.

\paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, data on which computations are applied among available processors (or cores). Each processor then applies the same progam or instruction onto its subpart of data. Moreover, if a neighborhood information is needed from another processor communications or synchronizations are performed.

In the domain of numerical simulations, this technique is most of the time called a domain decomposition. This parallelization technique produce efficient programs up to thousands processors or cores, but on certain conditions. First, each subpart of data has to be big enough to overlap communication time. Second, the partitioning of data has to be balanced among processors. Thus, if this parallelization technique is clearly adequate to structured meshes, easy to balance, it is not for unstructured meshes or irregular structures where the amount of work is not heterogenous.

\paragraph{Task parallelism} Another well known parallelization technique is to identify in a program the different tasks and which one are independant and can be launched concurrently. Most of the time, such a prallelization technique create a dependency directed acyclic graph (dependency dag) from a set of ordered tasks. Dependencies are found from read/write information for each task. Actually, if a task $i$ write a data $a$ and if a task $j$ read $a$, then $i$ has to be finished before $j$ is performed. From a dependency graph two different solutions are available. First, a static schedule of tasks is built, which could be a good solution if the tasks are regular, or use a dynamic scheduler to dynamically decide at runtime which task is executed on which processor.

\paragraph{Hybrid parallelism} Finally, it is also possible to combine both those parallelization techniques to get what is called an hybrid parallelization. The interest of an hybrid parallelization is to bring another source of parallelism if limits of a given technique are reach.

%----------------------------------------
\subsection{Component models}
%----------------------------------------
Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities called components. Component models improves code re-use by a separation of concerns, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}. In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.

In the rest of this paper will be used concepts of the component model $L^2C$~\cite{l2c}. $L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world. As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses). 
A \emph{port} in $L^2C$ is an entity embbedded in the component which makes possible an \emph{assembly} of components. An assembly of components is a way to actually connect components together and to produce a complete application, as a set of components and their interactions. A provided service inside a component is associated to a \emph{provide} port, while a required service is associated to a \emph{use} port. It is also possible to group more than one required service into a single port, as a list, called a \emph{use-multiple} port. As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle on the left of the component, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it, both on the right of the component.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$Comp_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=1.5cm of u] (p1) {};
   \node[component,right=1.5cm of p1] (C1) {$Comp_1$};
   \node[use,right=1.5cm of C1] (um) {$m$};

  \path[-]
    (p) edge node {} (C)
    (C.east) edge node {} (u)
    (C1)	edge node {} (um)
    (p1) edge node {} (C1);
\end{tikzpicture}
\caption{Two components, one with a provide, use and use-multiple ports, the second with a single provide port}
\label{fig:ports}
\end{center}
\end{figure}

In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}. As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$Comp_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$Comp_1$};
   \node[use,right=1.5 of C1] (um) {$m$};

  \path[-]
    (p) edge node {} (C)
    (C) edge node {} (u)
    (C1)	edge node {} (um)
    (u) edge node {} (C1);
\end{tikzpicture}
\caption{Component assembly of Figure~\ref{fig:ports}}
\label{fig:assembly}
\end{center}
\end{figure}

With its minimal set of ports, $L^2C$ is a light component model which does not introduce overheads at runtime, and which results in a HPC-oriented component model. Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a distribution of component on distributed architectures is managed. Actually, in CCA distant components (on different processors or machines) have to be managed manually inside components using MPI~\cite{Graham2009MSE} (Message Passing Interfaces), while $L^2C$ proposes a specific MPI port to link distant components and ease their communications. Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside the assembly, which increase separation of concerns and code reuse for parallel applications.

>>>>

% !!!!!!!!!!!! a mettre en plus court autre part !!!!!!!!!!!!!!!
%----------------------------------------
% \subsection{Parallelization techniques}
% \label{sect:parallel}
%----------------------------------------
% Three classical parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism. Those parallelization techniques are independent from the actual parallel hardware used.

% \paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, data on which computations are applied among available processors (or cores). Each processor then applies the same progam or instruction onto its subpart of data. Moreover, if a neighborhood information is needed from another processor communications or synchronizations are performed.

% In the domain of numerical simulations, this technique is most of the time called a domain decomposition. This parallelization technique produce efficient programs up to thousands processors or cores, but on certain conditions. First, each subpart of data has to be big enough to overlap communication time. Second, the partitioning of data has to be balanced among processors. Thus, if this parallelization technique is clearly adequate to structured meshes, easy to balance, it is not for unstructured meshes or irregular structures where the amount of work is not heterogenous.

% \paragraph{Task parallelism} Another well known parallelization technique is to identify in a program the different tasks and which one are independant and can be launched concurrently. Most of the time, such a prallelization technique create a dependency directed acyclic graph (dependency dag) from a set of ordered tasks. Dependencies are found from read/write information for each task. Actually, if a task $i$ write a data $a$ and if a task $j$ read $a$, then $i$ has to be finished before $j$ is performed. From a dependency graph two different solutions are available. First, a static schedule of tasks is built, which could be a good solution if the tasks are regular, or use a dynamic scheduler to dynamically decide at runtime which task is executed on which processor.

% \paragraph{Hybrid parallelism} Finally, it is also possible to combine both those parallelization techniques to get what is called an hybrid parallelization. The interest of an hybrid parallelization is to bring another source of parallelism if limits of a given technique are reach.

### 2015-07-28 WOLFHPC15/article/introduction.tex
wolfhpc15 23_07
fix compiler 28_07

\begin{itemize}
\item DSL interest: simplified HPC programming + perf
\item DSL limits: Lots of DSLs difficult to write, to maintain, to make portable etc.
\item Components introduction, this paper takes a first example of DSL to components
\item The domain targeted is multi-stencil simulations
\item Contributions: DSL to components try + agnostic coarse grain DSL for multi-stencil programs
\item Outline
\end{itemize}

>>>>

\begin{itemize}
\item DSL interest: simplified HPC programming + perf
\item A DSL for multi-stencil simulations + motivation fluid dynamics + bolzman + ...
\item DSL limits: Lots of DSLs difficult to write, to maintain, to make portable etc.
\item Components introduction, this paper takes a first example of DSL to components
\item Contributions: agnostic coarse grain DSL for multi-stencil programs + DSL to components try
\item Outline
\end{itemize}
